---
title: "Testing Missigness distribution"
output: html_document
date: "2023-12-14"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

# Testing if missigness is random using Hypergeometric distribution

Test if missing values are distributed randomly among two groups using 

Probability of observing 0-8 intensities in a condition of 8 samples,
- when 8 non-missing values were observed in all 16 samples - black line
- when 9 non-missing values were observed in 16 samples - red line
- If 7 non-missing values were observed in 16 samples - green line


```{r}
m <- 8
n <- 16 - 8
k <- 8
plot(0:20 , dhyper(0:20,m,n,k), type = 'l')
k <- 9
lines(0:20 , dhyper(0:20,m,n,k), type = 'l', col = 2)
k <- 7
lines(0:20 , dhyper(0:20,m,n,k), type = 'l', col = 3)

k <- 4
lines(0:20 , dhyper(0:20,m,n,k), type = 'l', col = "pink")


```


```{r}
m <- 8
n <- 16-8

plot(0:8 , phyper(0:8,m,n,8), type = 'l')
lines(0:8 , phyper(0:8,m,n,9), type = 'l', col = 2)
lines(0:8 , phyper(0:8,m,n,7), type = 'l', col = 3)
lines(0:8 , phyper(0:8,m,n,0), type = 'l', col = 3)

```

```{r}

testDistributionOfPresent <- function(observedA ,observedB, samplesA, samplesB  ){
 totalObserved = observedA + observedB
 nrA <- ifelse(observedA <= observedB, observedA, observedB)
 res <- phyper(nrA, samplesA, samplesB, totalObserved)
 return(res)
}

testDistributionOfPresent(2,2,8,8)
testDistributionOfPresent(4,4,8,8)
testDistributionOfPresent(1,1,8,8)
testDistributionOfPresent(7,7,8,8)
testDistributionOfPresent(7,7,21,30)
testDistributionOfPresent(10,10,21,30)
testDistributionOfPresent(2,20,21,30)

```

Simulate

```{r}
condA <- 8
condB <- 8
observedA <- sample(0:8, 20000, replace = TRUE)
observedB <- sample(0:8, 20000, replace = TRUE)
xb <- data.frame(observedA = observedA, observedB = observedB, condA = 8, condB = 8)

xb$condA <- 8
xb$condB <- 8
xb$pvals <- testDistributionOfPresent(observedA = xb$observedA, observedB = xb$observedB, samplesA = xb$condA, samplesB = xb$condB)

```




# Odds Ratio

$$
\begin{array}{|r|cc|}
\hline
& {\text { Diseased }} & {\text { Healthy }} \\
\hline
\text { Exposed } & 20 & 380 \\
\text { Not exposed } & 6 & 594 \\
\hline
\end{array}
$$


$$ \text{Relative risk} = \frac{D_E/(D_E + H_E)}{D_N/(D_N + H_N)}= \frac{D_E/V_E}{D_N/V_N}= \frac{20/400}{6/600}= \frac{.05}{.01} = 5\,.$$
The odds ratio is different.  The ''odds'' of getting the disease if exposed is $$D_E/H_E=20/380\approx .052 ,$$and the odds if not exposed is  $$D_N/H_N = 6/594 \approx .010 \,.$$ The odds ratio is the ratio of the two,

$$\text{Odds ratio} = \frac{D_E/H_E}{D_N/H_N}=   \frac{20/380}{6/594}\approx   \frac{.052}{.010} = 5.2\,.$$


```{r}
relativeRisk <- function(observedA, observedB, samplesA, samplesB) {
   rr <- (observedA/(observedA + observedB))/(samplesA/(samplesA+samplesB))
   return(rr)
}
xb$rr <- relativeRisk(observedA = xb$observedA, observedB = xb$observedB, samplesA = xb$condA, samplesB = xb$condB)

odsRatio <- function(observedA, observedB, samplesA, samplesB) {
   rr <- (observedA/observedB)/(samplesA/samplesB)
   return(rr)
}

xb$or <- odsRatio(observedA = xb$observedA, observedB = xb$observedB, samplesA = xb$condA, samplesB = xb$condB)
hist(xb$or, breaks = 100)
abline(v = 1, col = 2)
hist(log(xb$or))
```

Ein Volcano darf auch nicht fehlen

```{r}
plot(log(xb$or), -log10(xb$pvals))

xb$logor <- log(xb$or)

library(tidyverse)
xc <- xb |> filter(logor < 0.01, logor > -0.01)
plot(xc$observedA, xc$pvals)
plot(xc$observedB, xc$pvals)
plot(xc$observedA, xc$observedB)

```







# compare with fishers test

```{r}
TeaTasting <- matrix(c(1, 2, 3, 4),
       nrow = 2,
       dimnames = list(Guess = c("A", "B"),
                       Truth = c("A", "B")))
TeaTasting
fisher.test(TeaTasting, alternative = "greater")
```
