---
title: "Testing Missigness distribution"
output: html_document
date: "2023-12-14"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

```

# Testing if missigness is random using Hypergeometric distribution

Test if missing values are distributed randomly among two groups using 

Probability of observing 0-8 intensities in a condition of 8 samples,
- when 8 non-missing values were observed in all 16 samples - black line
- when 9 non-missing values were observed in 16 samples - red line
- If 7 non-missing values were observed in 16 samples - green line



```{r}

testDistributionOfPresent <- function(observedA ,observedB, samplesA, samplesB  ){
   totalObserved = observedA + observedB
   nrA <- ifelse(observedA <= observedB, observedA, observedB)
   res <- phyper(nrA, samplesA, samplesB, totalObserved)
   return(res)
}

testDistributionOfPresent(2,2,8,8)
testDistributionOfPresent(4,4,8,8)
testDistributionOfPresent(1,1,8,8)
testDistributionOfPresent(7,7,8,8)
testDistributionOfPresent(7,7,21,30)
testDistributionOfPresent(10,10,21,30)
testDistributionOfPresent(2,20,21,30)

```

Simulate dataset

```{r}
Nprot <- 1000
condA <- 8
condB <- 8
observedA <- sample(0:8, Nprot, replace = TRUE)
observedB <- sample(0:8, Nprot, replace = TRUE)
xb <- data.frame(observedA = observedA, observedB = observedB)

xb$samplesA <- condA
xb$samplesB <- condB
proteinID <- unique(stringi::stri_rand_strings(Nprot + 20,5))[1:Nprot]
xb$proteinID <- proteinID
xlater <- xb
```


```{r}
xb$pvals <- testDistributionOfPresent(
   observedA = xb$observedA,
   observedB = xb$observedB,
   samplesA = xb$samplesA,
   samplesB = xb$samplesB)

```




# Odds Ratio

$$
\begin{array}{|r|cc|}
\hline
& {\text { Diseased }} & {\text { Healthy }} \\
\hline
\text { Exposed } & 20 & 380 \\
\text { Not exposed } & 6 & 594 \\
\hline
\end{array}
$$


$$ \text{Relative risk} = \frac{D_E/(D_E + H_E)}{D_N/(D_N + H_N)}= \frac{D_E/V_E}{D_N/V_N}= \frac{20/400}{6/600}= \frac{.05}{.01} = 5\,.$$
The odds ratio is different.  The ''odds'' of getting the disease if exposed is $$D_E/H_E=20/380\approx .052 ,$$and the odds if not exposed is  $$D_N/H_N = 6/594 \approx .010 \,.$$ The odds ratio is the ratio of the two,

$$\text{Odds ratio} = \frac{D_E/H_E}{D_N/H_N}=   \frac{20/380}{6/594}\approx   \frac{.052}{.010} = 5.2\,.$$


```{r}
relativeRisk <- function(observedA, observedB, samplesA, samplesB) {
   rr <- (observedA/(observedA + observedB))/(samplesA/(samplesA+samplesB))
   return(rr)
}
xb$rr <- relativeRisk(observedA = xb$observedA,
                      observedB = xb$observedB,
                      samplesA = xb$samplesA, samplesB = xb$samplesB)

odsRatio <- function(observedA, observedB, samplesA, samplesB) {
   rr <- (observedA/observedB)/(samplesA/samplesB)
   return(rr)
}

xb$or <- odsRatio(observedA = xb$observedA, observedB = xb$observedB, samplesA = xb$samplesA, samplesB = xb$samplesB)
hist(xb$or, breaks = 100)
abline(v = 1, col = 2)
hist(log(xb$or))

```

Volcano is also needed:

```{r}
plot(log2(xb$or), -log10(xb$pvals))

```



# compare with fishers test

```{r}
# Define your data
group_A <- c(1, 1, 0, 0, 0)
group_B <- c(1, 1, 1, 1, 1)

group_A <- c(1, 1, 1, 1, 1)
group_B <- c(1, 1, 1, 1, 1)

group_A <- c(1, 1, 0,0,0)
group_B <- c(1, 1, 0,0,0)

# Perform Fisher's exact test
fisher_result <- fisher.test(
   matrix(c(sum(group_A == 1), sum(group_A == 0), 
            sum(group_B == 1), sum(group_B == 0)), nrow = 2))



```

# collect all the steps in a function

```{r}

computeForMissing <- function(xb,
                              observedA = "observedA",
                              observedB = "observedB",
                              samplesA = "samplesA",
                              samplesB = "samplesB"
) {
   relativeRisk <- function(observedA, observedB, samplesA, samplesB) {
      rr <- (observedA/(observedA + observedB))/(samplesA/(samplesA + samplesB))
      return(rr)
   }
   odsRatio <- function(observedA, observedB, samplesA, samplesB) {
      rr <- (observedA/observedB)/(samplesA/samplesB)
      return(rr)
   }
   apply_fischer <- function(proteinID,observedA, observedB, samplesA, samplesB){
      mat <- matrix(c(observedA, samplesA - observedA, 
                      observedB,samplesB - observedB), nrow = 2)
      fisher_result <- fisher.test(mat)
      return(data.frame(proteinID =  proteinID,
                        p_value = fisher_result$p.value,
                        OdsRatio = (fisher_result$estimate),
                        conf.lower = (fisher_result$conf.int[1]),
                        conf.higher = (fisher_result$conf.int[2]))
      )
   }
   
   xb$OdsRatioM <- odsRatio(
      observedA = xb[["observedA"]],
      observedB = xb[["observedB"]],
      samplesA = xb[["samplesA"]],
      samplesB = xb[["samplesB"]])
   xb$relativeRiskM <- relativeRisk(
      observedA = xb[["observedA"]],
      observedB = xb[["observedB"]],
      samplesA = xb[["samplesA"]],
      samplesB = xb[["samplesB"]])
   
   res <- vector(mode = "list", length(nrow(xb)))
   
   for (i in seq(nrow(xb))) {
      res[[i]] <- apply_fischer(
         xb[["proteinID"]][i],
         xb[["observedA"]][i],
         xb[["observedB"]][i],
         xb[["samplesA"]][i],
         xb[["samplesB"]][i] )
   }
   
   result <- dplyr::bind_rows(res)
   xx <- dplyr::inner_join(xb , result, by = c("proteinID" = "proteinID"))
   return(xx)
}

res <- computeForMissing(xlater)
plot(log2(res$OdsRatioM), -log10(res$p_value))

```
