---
title: "Simulate Protein level Data"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Model 1

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)

```


For proteins:
- the proteins have a FC either equal 1, 0. or -1, 10% have 1 80% have 0 and 10% have -1.

What we however are measuring are peptide spectrum matches. Let's assume we observing peptides.

For peptides:

- assume that number of peptides per protein follows a geometric distribution, $N_{pep} \sim Geo(p)$
- the fold change of peptides is the same as the fold change of the protein
  $$FC_{pep} \sim N(FC_{pro}, \sigma_{FC_{P}})$$
- Let's assume that the abundance of the peptide does on depend on protein abundance. The distribution of peptide intensities is $I_{pep} \sim LogNormal(\mu,\sigma) $


```{r}
peptideAbundances <- sim_data()
```


# Analyse simulated data with prolfqua

```{r}
library(prolfqua)

atable <- AnalysisTableAnnotation$new()
atable$sampleName = "sample"
atable$factors["group_"] = "group"
atable$hierarchy[["protein_Id"]] = "proteinID"
atable$hierarchy[["peptide_Id"]] = "peptideID"
atable$set_response("abundance")

config <- AnalysisConfiguration$new(atable)
adata <- setup_analysis(peptideAbundances, config)

lfqdata <- prolfqua::LFQData$new(adata, config)
lfqdata$is_transformed(TRUE)

lfqdata$remove_small_intensities(threshold = 1)
lfqdata$filter_proteins_by_peptide_count()

lfqdata$factors()
```

```{r}
pl <- lfqdata$get_Plotter()
lfqdata$hierarchy_counts()
lfqdata$config$table$hierarchy_keys_depth()

pl$heatmap()
```

```{r}
pl$intensity_distribution_density()
```


# Fit peptide model

```{r}
formula_Condition <-  strategy_lm("abundance ~ group_")
lfqdata$config$table$hierarchyDepth  <- 2

# specify model definition
modelName  <- "Model"
Contrasts <- c("B_over_A" = "group_B - group_A")
lfqdata$subject_Id()

mod <- prolfqua::build_model(
  lfqdata,
  formula_Condition)
aovtable <- mod$get_anova()


head(aovtable)
xx <- aovtable |> dplyr::filter(FDR < 0.5)
signif <- lfqdata$get_copy()
signif$data <- signif$data |> dplyr::filter(protein_Id %in% xx$protein_Id)
hmSig <- signif$get_Plotter()$heatmap()
hmSig
```

# Aggregate data

```{r}
lfqdata$config$table$hierarchyDepth <- 1
ag <- lfqdata$get_Aggregator()
ag$medpolish()
protData <- ag$lfq_agg
```

```{r}
protData$response()
formula_Condition <-  strategy_lm("medpolish ~ group_")

mod <- prolfqua::build_model(
  protData,
  formula_Condition)

contr <- prolfqua::Contrasts$new(mod, Contrasts)
v1 <- contr$get_Plotter()$volcano()
v1$FDR
ctr <- contr$get_contrasts()
U <- ctr |> filter(grepl("_U$", protein_Id)) 
hist(U$diff)
hist(U$FDR)
D <- ctr |> filter(grepl("_D$", protein_Id)) 
hist(D$diff)
hist(D$FDR)

```







